\documentclass[11pt]{article}
\usepackage[fleqn]{amsmath}\textwidth 6.25in
\oddsidemargin -0.25in
%\evensidemargin -0.5in
\topmargin -0.5in
\textheight 9.00in

\begin{document}

%\tracingcommands=1
\newlength{\hW} % heading box width
\settowidth{\hW}{\bf wvs-005r1}
%\settowidth{\hW}{Page \pageref{lastpage}\ of \pageref{lastpage}}
\makeatletter
\def\@biblabel#1{#1.}
\newcommand{\ps@twolines}{%
  \renewcommand{\@oddhead}{%
    1 May 2000\hfill\parbox[t]{\hW}{{\bf wvs-005r1}\newline
                          Page \thepage\ of \pageref{lastpage}}}%
\renewcommand{\@evenhead}{}%
\renewcommand{\@oddfoot}{}%
\renewcommand{\@evenfoot}{}%
}%
\makeatother
\pagestyle{twolines}

\vspace{-10pt}
\begin{tabbing}
\phantom{References: }\= \\
To: \>Dave, Nathaniel, Joe\\
Subject: \>What would Fred Krogh do for the Level 2 Program?\\
From: \>Van Snyder\\
\end{tabbing}

\parindent 0pt \parskip 3pt
\vspace{-20pt}

The mathematical problem immanent in the level 2 retrieval is a nonlinear
least-squares problem.  Nonlinear least-squares problems can be solved by
iterative methods based on local linearization; they require to solve
linear problems at each iteration.  In the case of the level 2 problem,
the linear problem that results has an unusual structure, the exploiting
of which could result in substantial computational economy.

Fred Krogh has an unique combination of knowledge, skills and talents in
computational mathematics and mathematical software.  In particular, he
has written two software packages to solve nonlinear least-squares
problems.  One of these packages is used by the trajectory planning
group; it routinely and efficiently solves problems that no other package
is able to solve at all.

Fred also has a deep understanding of the relation between linear
algebra, software to solve problems in linear algebra, and the solution
of nonlinear least-squares problems.  In particular,  Fred has experience
in developing software to solve linear algebra problems having unusual
structure, and incorporating such into software to solve nonlinear
least-squares problems.

Fred has already given me substantial advice on methods and software to
solve nonlinear least-squares problems, and software and algorithms to
solve the linear problems that arise in solving nonlinear least-squares
problems.  In particular, he has given me software, that already has many
of the features we will need in the level 2 program, to solve nonlinear
least-squares problems.  I would find his advice to be useful in the
development of the level 2 program.  I therefore propose to institute a
consulting contract with Fred, for 17 days of work (20 hours per month
for the remainder of the fiscal year), at \$47.50 per hour.  This would
cost \$6460.00.  I expect that exploiting his knowledge, skills and
talents could save several work-months of effort, and result in software
of superior reliability, accuracy and performance.

Pursuant to this consulting agreement, Fred would provide mathematical
development, advice on mathematical approaches, advice on computational
approaches, advice on specialized software organization strategies,
advice on selection of existing software, and advice on modification of
existing software.

Fred is a JPL retiree, and already has a consulting agreement with JPL. 
I understand that the paperwork to add funds and work authorization to an
existing contract is substantially simpler than creating a new consulting
agreement.

\label{lastpage}
\end{document}
% $Id$
